Effectful Semantics for Substructural SSA

### 2023-11-06 pldi24/paper.tex
Tap...

@@

@@

>>>>

\label{sec:semantics}

@@

In this section, we will build on the bare-bones models given in Section \ref{sec:semantics} to develop more sophisticated models capable of underpinning \isotopessa dialects supporting features such as weak memory and separation logic.

@@

Building on the ideas in \cite{mellies-ftrs}, we will show how a simple separation logic based on \cite{reynolds-separation-2002} slightly adapted to the categorical setting gives us a very rich \isotopessa model supporting a basic notion of refinement types, while at the same time giving us a nontrivial notion of nonlinear types.

We'll start by considering the category \(\ms{HRel}\), which we will take to have morphisms of the form \(A \to \ms{Heap} \to \mc{P}^+(A \times \ms{Heap} \times \ms{String}^* + \ms{String}^\omega)\), which is the result of applying the state-transformer monad with state \(\ms{Heap}\) to the trace monad with the infinite monoid of lists of strings described in the previous section.

% \TODO{by the above, this is an \isotopessa model w/ nice properties; moving on!}

% We now define a separation logic with predicates having the following grammar:
% \begin{grammar}
%   <\(\varphi, \psi\)> ::= 
%     \(\varnothing\) 
%     \;|\; \(e \mapsto a\) 
%     \;|\; \(\varphi * \psi\)
%     \;|\; \(\forall x: A, \varphi\)
%     \;|\; \(\exists x: A, \varphi\)
%     \;|\; \(\varphi \land \psi\)
%     \;|\; \(\varphi \lor \psi\)
%     \;|\; \(\varphi \Rightarrow \psi\)
%     \;|\; \(\top\)
%     \;|\; \(\bot\)
% \end{grammar}
% where \(e, a\) denote regular \isotopessa expressions and \(A, B\) regular \isotopessa types.

% We can now do something somewhat tricky with \isotopessa dialects:
% \TODO{truly pure expressions \(\implies\) semantics for expressions}
% \TODO{giving us a semantics for separation logic stuff}
% \TODO{now we define the category \(\ms{Sep}\) w/ objects contexts + predicates}
% \TODO{tensor product, niceties \(\implies\) category, yey}
% \TODO{sliding}
% \TODO{therefore, substructural \textit{effectful} w/ nontrivial linear objects}
